Appendix

### 2023-10-03 manuscript/_appendix/_syllable-frequency.qmd
Ignore .tiff, .eps, and .svg

items <- items |> 
    mutate(xsampa = ipa::ipa(ipa)) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = purrr::map_int(syll, length),
           item = gsub("cat_|spa_", "", item)) 
```

@@

```{r tbl-items-table}
#| label: tbl-syll-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish
#| echo: false
#| warning: false
#| message: false
tbl_data <- items |> 
    distinct(te, item, language, meaning, xsampa, n_syll) |> 
    left_join(distinct(pool_tmp, te, language, item, label),
              by = join_by(te, item, language)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, xsampa, n_syll),
                id_cols = c(te, meaning),
                names_repair = janitor::make_clean_names,
                values_fn = first) |> 
    sample_n(15) |> 
    select(-c(te, item_catalan, item_spanish)) |> 
    relocate(meaning, matches("catalan"), matches("spanish"))

@@

tbl_data |> 
    kable(format = "latex", 
          digits = 3,
          booktabs = TRUE,
          col.names = col_names, 
          align = "lllcllc") 
```

@@

```{r fig-syll-number}
#| label: fig-syll-number
#| fig-cap: Distribution of the number of syllables in Catalan and Spanish
#| fig-width: 7
#| fig-height: 3
#| echo: false
#| warning: false
#| message: false
pool_tmp |> 
    group_by(language) |> 
    add_count(language, name = "n_lang") |> 
    ungroup() |> 
    count(language, n_lang, n_syll) |> 
    mutate(prop = n/n_lang) |> 
    ggplot(aes(n_syll, n)) +
    facet_wrap(~ language) +
    geom_col(colour = "white", fill = "grey70") +
    geom_text(aes(label = scales::percent(prop)),
              size = 3,
              position = position_nudge(y = 15),
              colour = "black") +
    labs(x = "Number of syllables",
         y = "Number of words",
         fill = "Language") +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
```

@@

syll |> 
    add_count(language, name = "n_lang") |> 
    mutate(language = paste0(language, " (n = ", n_lang, ")")) |> 
    ggplot(aes(freq_syll)) +
    facet_wrap(~language) +
    geom_histogram(colour = "white",
                   bins = 20,
                   fill = "grey70",
                   position = position_nudge(y = 0.1)) +
    geom_rug(alpha = 1/10) +
    labs(x = "Relative frequency (counts per million)",
         y = "Counts",
         fill = "Language",
         colour = "Language") +
    scale_x_continuous(trans = "log10",
                       labels = \(x) format(x, big.mark = ",")) +
    theme(legend.title = element_blank(),
          panel.grid = element_blank(),
          legend.position = "none")
```

To estimate the association between word-level syllabic frequency and cognateness, while controlling for the number of syllables in the word, as words are expected to necessarily increase the syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ [Levenshtein similarity between a word and its translation equivalent, @levenshtein1966binary] as predictors. We added translation equivalent-level random effects for the intercept and the main effect of $Syllables$ (some translation pairs had a different number of syllables in each language). We used a Gaussian distribution to model syllabic frequency scores after standardising this variable and the predictors. We used a weakly informative prior for all parameters involved in the model (see @eq-model-syllables for a formal equation of this model and its prior). We conducted statistical inference by evaluating the proportion of the 95% highest density interval (HDI) of the posterior posterior distribution of each coefficient that fell into the region of practical equivalence [ROPE, see the main manuscript for a more detailed explanation, @kruschke2018bayesian].

@@

```{r syllables-posterior}
#| label: syllables-posterior
#| echo: false
#| warning: false
#| message: false
rope_interval <- c(lower = -0.5, upper = 0.5)
# tidy predictor names

post_draws_list <- posterior_syllables_summary |> 
    split(posterior_syllables_summary$.variable)

@@

```{r tbl-syll-coefs}
#| label: tbl-syll-coefs
#| tbl-cap: "Posterior distribution of regression coefficients."
#| tbl-note: "Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
tbl_data <- posterior_syllables_summary |>
    select(.variable_name, .value, .lower, .upper, .rope) |>
    mutate(
        .variable_name = gsub("×", '$\\\\times$', .variable_name),
        .hdi = paste0("[", round(.lower, 3),
                      ", ", round(.upper, 3), "]"),
        .rope = sub("^(-?)0.", "\\1.",
                    sprintf("%.3f", .rope))) |> 
    select(.variable_name, .value, .lower, .upper, .hdi, .rope)

tbl_data |> 
    select(-c(.lower, .upper)) |> 
    knitr::kable(
        format = "latex", 
        escape = FALSE,
        col.names = c("",
                      "$\\beta$",
                      "95\\% HDI",
                      "$p(\\text{ROPE})$"),
        digits = 3,
        booktabs = TRUE,
        align = "lclr",
        linesep = ""
    ) |> 
    kableExtra::kable_styling()
```

@@

```{r fig-syll-marginal}
#| label: fig-syll-marginal
#| fig-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
#| fig-width: 7
#| fig-height: 5
#| out-width: 80%
#| echo: false
#| message: false
#| warning: false
expand_grid(lv_std = scale(seq(0, 1, 0.1),
                           center = mean(syllables_data$lv),
                           scale = sd(syllables_data$lv))[, 1],
            n_syll_std = scale(seq(1, 4, 1),
                               center = mean(syllables_data$n_syll),
                               scale = sd(syllables_data$n_syll))[, 1]) |> 
    tidybayes::add_epred_draws(model_fit_syllables,
                               re_formula = NA) |>
    tidybayes::median_hdci() |> 
    mutate(
        lv_std = (lv_std * sd(syllables_data$lv)) + mean(syllables_data$lv),
        n_syll_std = (n_syll_std * sd(syllables_data$n_syll)) + mean(syllables_data$n_syll),
        n_syll_std = as.factor(n_syll_std)) |> 
    ggplot(aes(lv_std, .epred, 
               shape = n_syll_std)) +
    geom_point(data = filter(syllables_data, n_syll < 5) |> 
                   mutate(n_syll_std = as.factor(n_syll)),
               aes(x = lv,
                   y = freq_syll,
                   shape = n_syll_std),
               alpha = 1/2) +
    # geom_ribbon(aes(ymin = .lower,
    #                 ymax = .upper),
    #             alpha = 1/2,
    #             colour = NA,
    #             show.legend = FALSE) +
    geom_line(aes(linetype = n_syll_std),
              linewidth = 0.5) +
    labs(x = "Cognateness (Levenshtein similarity)",
         y = "Syllabic frequency",
         linetype = "Number of syllables",
         shape = "Number of syllables",
         colour = "Number of syllables") +
    scale_y_continuous(n.breaks = 8) +
    theme(panel.grid = element_blank(),
          legend.position = "top")
```

>>>>

items <- items |> 
    mutate(xsampa = ipa::ipa(ipa),
           syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = purrr::map_int(syll, length),
           item = gsub("cat_|spa_", "", item)) 
```

@@

```{r tbl-items-table}
#| label: tbl-syll-items
#| tbl-cap: Sample of items included in the BVQ questionnaire and their syllabified SAMPA transcriptions in Catalan and Spanish.
#| echo: false
#| warning: false
#| message: false
tbl_data <- items |> 
    distinct(te, item, language, meaning, xsampa, n_syll) |> 
    left_join(distinct(pool_tmp, te, language, item, label),
              by = join_by(te, item, language)) |> 
    pivot_wider(names_from = language,
                values_from = c(item, label, xsampa, n_syll),
                id_cols = c(te, meaning),
                names_repair = janitor::make_clean_names,
                values_fn = first) |> 
    sample_n(15) |> 
    select(-c(te, item_catalan, item_spanish)) |> 
    relocate(meaning, matches("catalan"), matches("spanish"))

@@

tbl_data |> 
    kable(digits = 3,
          format = "latex",
          booktabs = TRUE,
          col.names = col_names, 
          align = "lllcllc") 
```

@@

```{r fig-syll-number}
#| label: fig-syll-number
#| fig-cap: Distribution of the number of syllables in Catalan and Spanish
#| fig-width: 7
#| fig-height: 3
#| echo: false
#| warning: false
#| message: false
pool_tmp |> 
    group_by(language) |> 
    add_count(language, name = "n_lang") |> 
    ungroup() |> 
    count(language, n_lang, n_syll) |> 
    mutate(prop = n/n_lang) |> 
    ggplot(aes(n_syll, n)) +
    facet_wrap(~ language) +
    geom_col(colour = "white", fill = "black") +
    geom_text(aes(label = scales::percent(prop)),
              size = 3,
              position = position_nudge(y = 15),
              colour = "black") +
    labs(x = "Number of syllables",
         y = "Number of words",
         fill = "Language") +
    theme(legend.position = "none",
          panel.grid = element_blank())
```

@@

syll |> 
    add_count(language, name = "n_lang") |> 
    mutate(language = paste0(language, " (n = ", n_lang, ")")) |> 
    ggplot(aes(freq_syll)) +
    facet_wrap(~language) +
    geom_histogram(colour = "white",
                   bins = 20,
                   fill = "black",
                   position = position_nudge(y = 0.1)) +
    geom_rug(alpha = 1/10) +
    labs(x = "Relative frequency (counts per million)",
         y = "Counts",
         fill = "Language",
         colour = "Language") +
    scale_x_continuous(trans = "log10",
                       labels = \(x) format(x, big.mark = ",")) +
    theme(legend.title = element_blank(),
          panel.grid = element_blank(),
          legend.position = "none")
```

To estimate the association between word-level syllabic frequency and cognateness, while controlling for the number of syllables in the word, (an increment in the number of syllables necessarily increases the summed syllabic frequency of the word), we fit a multilevel, Bayesian linear regression model with syllabic frequency (the sum of the syllabic frequency of the syllables in a word) as response variable, and the main effect of the number of syllables ($Syllables$) and $Cognateness$ [Levenshtein similarity between a word and its translation equivalent, @levenshtein1966binary] as predictors. We added translation equivalent-level random effects for the intercept and the main effect of $Syllables$ (some translation pairs had a different number of syllables in each language). We used a Gaussian distribution to model syllabic frequency scores after standardising this variable and the predictors. We used a weakly informative prior for all parameters involved in the model (see @eq-model-syllables for a formal equation of this model and its prior). We conducted statistical inference by evaluating the proportion of the 95% highest density interval (HDI) of the posterior posterior distribution of each coefficient that fell into the region of practical equivalence [ROPE, see the main manuscript for a more detailed explanation, @kruschke2018bayesian].

@@

```{r syllables-posterior}
#| label: syllables-posterior
#| echo: false
#| warning: false
#| message: false
rope_interval <- c(lower = -0.5, upper = 0.5)

# tidy up predictor names
post_draws_list <- posterior_syllables_summary |> 
    split(posterior_syllables_summary$.variable)

@@

```{r tbl-syll-coefs}
#| label: tbl-syll-coefs
#| tbl-cap: "Posterior distribution of regression coefficients."
#| tbl-note: "Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
tbl_data <- posterior_syllables_summary |>
    select(.variable_name, .value, .lower, .upper, .rope) |>
    mutate(
        .variable_name = gsub("×", '$\\\\times$', .variable_name),
        .hdi = paste0("[", round(.lower, 3),
                      ", ", round(.upper, 3), "]"),
        .rope = sub("^(-?)0.", "\\1.",
                    sprintf("%.3f", .rope)),
        .rope = ifelse(is.na(.rope), "-", .rope)) |> 
    select(.variable_name, .value, .lower, .upper, .hdi, .rope)

col_names <- c("", "$\\beta$", "95\\% HDI", "$p(\\text{ROPE})$")

tbl_data |> 
    select(-c(.lower, .upper)) |> 
    knitr::kable(
        format = "latex",
        escape = FALSE,
        col.names = col_names,
        digits = 3,
        booktabs = TRUE,
        align = "lclr",
        linesep = ""
    ) |> 
    kableExtra::kable_styling() |> 
    add_footnote("$\\beta$: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution. $p(\\text{ROPE})$: overlap between the 95\\% HDI and the ROPE, indicating the posterior probability that the true value of the coefficient is equivalent to zero.", notation = "none")
```

@@

```{r fig-syll-marginal}
#| label: fig-syll-marginal
#| fig-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
#| fig-width: 9
#| fig-height: 2.5
#| out-width: 80%
#| echo: false
#| message: false
#| warning: false
syll_plot_data <- expand_grid(
    lv_std = scale(seq(0, 1, 0.1),
                   center = mean(syllables_data$lv),
                   scale = sd(syllables_data$lv))[, 1],
    n_syll_std = scale(seq(1, 4, 1),
                       center = mean(syllables_data$n_syll),
                       scale = sd(syllables_data$n_syll))[, 1]) |> 
    tidybayes::add_epred_draws(model_fit_syllables,
                               re_formula = NA) |>
    tidybayes::median_hdci() |> 
    mutate(
        lv_std = (lv_std * sd(syllables_data$lv)) + mean(syllables_data$lv),
        n_syll_std = (n_syll_std * sd(syllables_data$n_syll)) + mean(syllables_data$n_syll),
        n_syll_std = as.factor(n_syll_std)
    )

ggplot(syll_plot_data, aes(lv_std, .epred)) +
    facet_grid(~n_syll_std) +
    geom_point(data = filter(syllables_data, n_syll < 5) |>
                   mutate(n_syll_std = as.factor(n_syll)),
               aes(x = lv,
                   y = freq_syll),
               alpha = 1/5,
               shape = 1) +
    geom_ribbon(aes(ymin = .lower,
                    ymax = .upper),
                alpha = 1/2,
                colour = NA,
                show.legend = FALSE) +
    geom_line(linewidth = 0.5) +
    labs(x = "Cognateness (Levenshtein similarity)",
         y = "Syllabic frequency",
         linetype = "Number of syllables",
         shape = "Number of syllables",
         colour = "Number of syllables") +
    scale_y_continuous(n.breaks = 8) +
    theme(panel.grid = element_blank(),
          legend.position = "top")

suppressMessages({
    invisible({
        devices <- c("png", "pdf", "svg", "tiff", "eps")
        paths <- paste("_assets/img/Figure_6_SuppInfo", devices, sep = ".")
        purrr::map(paths, \(x) ggsave(x, dpi = 1e3L, width = 9, height = 2.5, units = "in"))
    })
})
```

### 2023-10-03 manuscript/appendix.qmd
Ignore .tiff, .eps, and .svg

my_theme <- theme_minimal() +
    theme(
        panel.grid = element_line(
            colour = "grey",
            linetype = "dotted"
        ),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 12, colour = "black"),
        axis.text = element_text(colour = "black")
    )

theme_set(my_theme)
```

## Appendix A: model diagnostics {.appendix .unnumbered}

{{< include _appendix/_model-diagnostics.qmd >}}

>>>>

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(
        colour = "grey",
        linetype = "dotted"
    ),
    axis.line = element_line(colour = "black"),
    text = element_text(size = 12, colour = "black"),
    axis.text = element_text(colour = "black")
    )

theme_set(my_theme)

save_image <- function(file, ...) {
    suppressWarnings({
        invisible({
            devices <- c("png", "pdf", "svg", "tiff", "eps")
            paths <- file.path(paste(file, devices, sep = "."))
            purrr::map(paths,
                       \(x) ggsave(x, ...),
                       .progress = TRUE)
        })
    })
}

post_draws_list <- model_summary |>
    mutate(across(
        .value:.upper,
        \(x) ifelse(grepl("Intercept", .type),
                    plogis(x), x / 4
        )
    )) |>
    split(model_summary$.variable)
```

## Appendix A: model details {.appendix .unnumbered}

{{< include _appendix/_model-details.qmd >}}
